Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics

### 2024-08-23 Thesis/04-chap4.qmd

# Software for Automated Residual Plot Assessment: autovi and autovi.web

@@

Region 2 of @fig-autovi-web is a panel for data uploading and CSV type selection. Clicking the "upload CSV" button opens a window where the user can select a file from their local system. The data status displayed above the button provides information about the number of rows and columns in the current dataset. Additionally, there are two example datasets available beneath the "upload CSV" button: one is a lineup example using a CSV file with three columns, and the other is a single plot example using a CSV file with two columns. More details about these example datasets are be discussed in @sec-autovi-web-example.

@@

### Usage, from the user perspective

The user would follow the order of the numbered steps displayed in the main user interface as shown in @fig-autovi-web, <!--Since the design details have already been covered in @sec-autovi-web-design, we will not repeat the functionality of each panel in this section.-->
as follows:

- Step 1: uploading a CSV file by clicking the "upload CSV" button. A window will pop up, allowing the user to select the CSV file. The CSV file should contain at least two columns representing fitted values and residuals of a residual plot. If a lineup needs to be evaluated, the CSV file should contain an additional column for labels of residual plot.  
- In Step 2, the user specifies the CSV type based on the uploaded file, choosing between a single residual plot or a lineup of multiple residual plots. 
- Step 3 involves selecting the appropriate columns for fitted values, residuals, and, optionally, labels of residual plots, from the dataset. 
- In Step 4, the user sets the number of bootstrapped draws needed for the analysis; more draws provide a better estimate of the bootstrapped distribution but result in slower computation. The user should also set the seed to control the simulation. If the dataset contains a lineup and the true residual plot is known, select its label. 
- Finally, clicking the play button will initiate the analysis, and the user can review the result panels.

### Example {#sec-autovi-web-example}

>>>>

# Software for Automated Residual Plot Assessment: autovi and autovi.web {#sec-third-paper}

@@

Region 2 of @fig-autovi-web is a panel for data uploading and CSV type selection. Clicking the "upload CSV" button opens a window where the user can select a file from their local system. The data status displayed above the button provides information about the number of rows and columns in the current dataset. Additionally, there are two example datasets available beneath the "upload CSV" button: one is a lineup example using a CSV file with three columns, and the other is a single plot example using a CSV file with two columns. More details about these example datasets are be discussed in @sec-autovi-web-workflow.

@@

### Workflow {#sec-autovi-web-workflow}

The workflow of `autovi.web` is designed to be straightforward, with numbered steps displayed in each panel shown in @fig-autovi-web. Thre are two example datasets provided by the web application, as mentioned in @sec-autovi-web-design. The single residual plot example uses the `dino` dataset from the R package `datasauRus` [@datasaurus]. The lineup example uses residuals from a simulated regression model that has a non-linearity issue. We will walk through the lineup example to further demonstrate the workflow of the web application.

As shown in @fig-autovi-web-workflow-example, to use the lineup example data, click the "Use Lineup Example" button. The data status will then update to show the number of rows and columns in the dataset, and the CSV type will automatically be selected to the correct option. Since the example dataset follows the variable naming conventions assumed by the web application, the columns for fitted values, residuals, and labels of residual plots are set automatically. If the user is working with a custom dataset, these options must be set accordingly. Regardless of the dataset, the user must manually select the label for the true residual plot, as the web application allows assessments without this label. The next step is to click the play button to start the assessment.

In the result panels shown in [@fig-autovi-web-workflow-example]D, the first row of the table is the most crucial to check, as it provides the visual signal strength and the rank of the true residual plot among the other plots. The summary text beneath the table provides the $p$-value, which can be used for quick decision-making. The lineup is for manual inspection, and the user should see if the true residual plot is visually distinguishable from the other plots, to confirm if the model violation is serious. The density plot in [@fig-autovi-web-workflow-example]E offers a more robust result, allowing the user to compare the distribution of bootstrapped visual signal strength with the distribution of null visual signal strength. Finally, the grayscale attention map can be used to check if the target visual features, like the non-linearity present in the lineup example, are captured by the computer vision model, ensuring the quality of the assessment.

```{r fig-autovi-web-workflow-example, out.width = "100%", fig.cap = 'The workflow for `autovi.web` using the lineup example dataset proceeds as follows: (A) To begin, the user clicks the "Use Lineup Example" button to load the example dataset, during which the data status and CSV type will be automatically updated. (B) The target columns are selected automatically, though the user must manually select the label for the true residual plot, as the web application permits assessment without this label. (C) The user initiates the assessment by clicking the run button. (D) The visual signal strength of the true residual plot is displayed in the first row of the table, with a summary text beneath the table providing the $p$-value to aid in decision-making. The lineup on the right allows for manual inspection. (E) The density plot helps verify if the bootstrapped distribution differs from the null distribution, while the attention map on the right offers insights into whether the computer vision model has captured the intended visual features of the true residual plot.'}
magick::image_read_pdf("figures/autovi_web_workflow.pdf")
```
